Appendix 1: sdmTMB statistical model description

### 2021-12-22 doc/appendix-model.Rmd
Develop appendix

---
title: "Appendix 1: sdmTMB statistical model description"
output:
  bookdown::pdf_document2:
    toc: false
number_sections: false
fig_caption: true
urlcolor: blue
highlight: "monochrome"
csl: mee.csl
bibliography: refs.bib
header-includes:
  \usepackage[left]{lineno}
  \usepackage{bm}
  \usepackage{amssymb}
---

@@

  "$\\sigma_\\alpha^2$", "IID random intercept variance",
  "$\\sigma_\\epsilon^2$", "Spatiotemporal random field marginal variance",
  "$\\sigma_\\omega^2$", "Spatial random field marginal variance",
  "$\\sigma_\\zeta^2$", "Spatially varying coefficient random field marginal variance",
  "$\\rho$", "Correlation between random fields in subsequent time steps"
)
knitr::kable(symbols, caption = "Symbol notation and descriptions.", escape = FALSE, format = "latex", booktabs = TRUE,
   linesep = c(
    rep('', 7), # y, mean etc.
    '\\addlinespace',
    rep('', 7), # fields
    '\\addlinespace',
    rep('', 5), # covariance stuff
    '\\addlinespace',
    rep('', 2), # bayes
    '\\addlinespace'
  ))
```

@@

where $\tau^2$ controls the spatial variance,
$\nu$ controls the smoothness,
$\Gamma$ represents the Gamma function,
$d_{jk}$ represents the distance between locations $s_j$ and $s_k$,
$K_\nu$ represents the modified Bessel function of the second kind,
and $\kappa$ controls the spatial correlation range.
The parameter $\nu$ is set to 1 to take advantage of the Stochastic Partial Differential Equation (SPDE) approximation to the GMRF to greatly increase computational efficiency [@lindgren2011].

## Projection '$\boldsymbol{A}$' matrix

>>>>

---
title: "Appendix 1: sdmTMB statistical model description"
output:
  bookdown::pdf_document2:
    toc: true
    number_sections: false
fig_caption: true
urlcolor: blue
highlight: "monochrome"
csl: mee.csl
bibliography: refs.bib
header-includes:
  \usepackage[left]{lineno}
  \usepackage{bm}
  \usepackage{amssymb}
---

@@

  "$\\sigma_\\alpha^2$", "IID random intercept variance",
  "$\\sigma_\\epsilon^2$", "Spatiotemporal random field marginal variance",
  "$\\sigma_\\omega^2$", "Spatial random field marginal variance",
  "$\\sigma_\\zeta^2$", "Spatially varying coefficient random field marginal variance",
  "$\\kappa_\\omega$", "Spatial decorrelation rate",
  "$\\kappa_\\epsilon$", "Spatiotemporal decorrelation rate",

  "$\\rho$", "Correlation between random fields in subsequent time steps",
  "$\\boldsymbol{A}$", "A sparse projection matrix to bilinearly interpolate between knot and data locations",
  "$\\boldsymbol{H}$", "A 2-parameter rotation matrix used to define anisotropy"
)
knitr::kable(symbols, caption = "Symbol notation and descriptions.", escape = FALSE, format = "latex", booktabs = TRUE,
   linesep = c(
    rep('', 7), # y, mean etc.
    '\\addlinespace',
    rep('', 7), # fields
    '\\addlinespace',
    rep('', 5), # covariance stuff
    '\\addlinespace',
    rep('', 5), # more covariance stuff
    '\\addlinespace',
    rep('', 99) # end
  ))
```

@@

where $\tau^2$ controls the spatial variance,
$\nu$ controls the smoothness,
$\Gamma$ represents the Gamma function,
$d_{jk}$ represents the distance between locations $s_j$ and $s_k$,
$K_\nu$ represents the modified Bessel function of the second kind,
and $\kappa$ controls the spatial correlation range (decorrelation rate).
The parameter $\nu$ is set to 1 to take advantage of the Stochastic Partial Differential Equation (SPDE) approximation to the GMRF to greatly increase computational efficiency [@lindgren2011].
Internally, the parameter $\kappa$ and $\tau$ are converted to range and marginal standard deviation $\sigma$ as $\textrm{range} = \sqrt{8} / \kappa$ and $\sigma = 1 / \sqrt{4 \pi \exp \left(2 \log(\tau) + 2 \log(\kappa) \right) }$.

In the case of a spatiotemporal model with both spatial and spatiotemporal fields, if `share_range` is `TRUE` in `sdmTMB()` (the default), then a single $\kappa$ and range are estimated with separate $\sigma_\omega$ and $\sigma_\epsilon$. This often makes sense since the data are often only weakly informative about $\kappa$.
If `share_range = FALSE`, then separate $\kappa_\omega$ and $\kappa_\epsilon$ are estimated.
The spatially varying coefficient field always shares $\kappa$ with the spatial random field.

## Projection $\boldsymbol{A}$ matrix
